---
title: "Week-10- Diagnostic Plots"
author: "Nishad Tupe"
date: "March 20, 2019"
output: html_document
Reference: https://data.library.virginia.edu/diagnostic-plots/
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## understanding diagnostic plot for linear regression analysis.



Along with residuals built in diagonostic plots for linear regression
that helps us how well the model explains the relationship within predictor and # response variable

```{r cars}
data(women)
```

## Load the dataset

```{r pressure, echo=FALSE}
fit = lm(weight ~ height,women)
```
# Set Par - To look at four plots at once rather than one by one
# Change the panel layout to 2 x 2
```{r}
par(mfrow=c(2,2)) 
```
```{r}
plot(fit)
```
# Change the panel layout to 1 x 1
```{r}
par(mfrow=c(1,1)) 
```

# Residuals vs Fitted - This plot shows if residuals have non-linear patterns
# Normal Q-Q - This plot shows if residuals are normally distributed. Do residuals follow a straight line well or do they deviate severely? It's good if residuals are lined well on the straight dashed line.
# Scale-Location - It's also called Spread-Location plot. This plot shows if residuals are spread equally along the ranges of predictors.  
# Residuals vs Leverage - This plot helps us to find influential cases (i.e., subjects) if any.





