---
title: "Tidy data"
author: "Nishad Tupe"
date: "2/6/2019"
output:
  html_document:
    df_print: paged
Reference: https://r4ds.had.co.nz/tidy-data.html
---

# Loading all the necessary libraries.


```{r}
library(tidyverse)
library(readr)
library(tidyr)
library(ggplot2)
library(dplyr)
```

# Look at data in table1
```{r}
table1
```

# Compute rate per 10,000

```{r}
table1 %>% 
  mutate(rate = cases / population * 10000)
```
# Compute the Cases per Year

```{r}
table1 %>% 
count(year, wt = cases)

```
# Visualize changes over time

```{r}
library(ggplot2)
ggplot(table1, aes(year, cases)) + 
  geom_line(aes(group = country), colour = "grey50") + 
  geom_point(aes(colour = country))
```

# Gathering
```{r}
table4a
```

# Using Gather
To tidy a dataset like this, we need to gather those columns into a new pair of variables. To describe that operation we need three parameters: 
The set of columns that represent values, not variables. In this example, those are the columns 1999 and 2000.
The name of the variable whose values form the column names. I call that the key, and here it is year.
The name of the variable whose values are spread over the cells. I call that value, and here it's the number of cases.
Together those parameters generate the call to gather()

```{r}
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
```
# Using Left_join()

To combine the tidied versions of table4a and table4b into a single tibble, we need to use dplyr::left_join()

```{r}
tidy4a <- table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
tidy4b <- table4b %>% 
  gather(`1999`, `2000`, key = "year", value = "population")
left_join(tidy4a, tidy4b)
```

# Spreading
Spreading is the opposite of gathering. You use it when an observation is scattered across multiple rows. For example, take table2: an observation is a country in a year, but each observation is spread across two rows.

```{r}
table2
```

To tidy this up, we first analyse the representation in similar way to gather(). This time, however, we only need two parameters:
The column that contains variable names, the key column. Here, it's type.
The column that contains values from multiple variables, the value column. Here it's count.
Once we've figured that out, we can use spread()

```{r}
table2 %>%
    spread(key = type, value = count)
```

# Separating and uniting

## separate
separate() pulls apart one column into multiple columns, by splitting wherever a separator character appears. 
```{r}
table3 %>% 
  separate(rate, into = c("cases", "population"))
```
## Unite

```{r}
table5 %>% 
  unite(new, century, year)
```

# Missing values

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr    = c(   1,    2,    3,    4,    2,    3,    4),
  return = c(1.88, 0.59, 0.35,   NA, 0.92, 0.17, 2.66)
)
```
There are two missing values in this dataset:
The return for the fourth quarter of 2015 is explicitly missing, because the cell where its value should be instead contains NA.
The return for the first quarter of 2016 is implicitly missing, because it simply does not appear in the dataset.

```{r}
stocks %>% 
  spread(year, return)
```

Because these explicit missing values may not be important in other representations of the data, you can set na.rm = TRUE in gather() to turn explicit missing values implicit:

```{r}
stocks %>% 
  spread(year, return) %>% 
  gather(year, return, `2015`:`2016`, na.rm = TRUE)
```
complete() takes a set of columns, and finds all unique combinations. It then ensures the original dataset contains all those values, filling in explicit NAs where necessary.

```{r}
stocks %>% 
  complete(year, qtr)
```

There's one other important tool that you should know for working with missing values. Sometimes when a data source has primarily been used for data entry, missing values indicate that the previous value should be carried forward

```{r}
treatment <- tribble(
  ~ person,           ~ treatment, ~response,
  "Derrick Whitmore", 1,           7,
  NA,                 2,           10,
  NA,                 3,           9,
  "Katherine Burke",  1,           4
)
```

You can fill in these missing values with fill(). It takes a set of columns where you want missing values to be replaced by the most recent non-missing value (sometimes called last observation carried forward).

```{r}
treatment %>% 
  fill(person)
```
