---
title: "Data Analysis"
author: "Nishad Tupe"
date: "March 20, 2019"
output: html_document
reference: http://daviddalpiaz.github.io/appliedstats/model-diagnostics.html#data-analysis-examples
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# How to Perform good dianostics using regression models.



## Building linear regression model and perform diagnostic.
## Use mtcars data with mpg as response and hp and am as predictors

```{r cars}
mpg_hp_add = lm(mpg ~ hp + am, data = mtcars)
plot(fitted(mpg_hp_add), resid(mpg_hp_add), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residual",
     main = "mtcars: Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)
```

## There are no patterns observed with constant variance.
## The Q-Q plot looks extremely good 
## The Shapiro-Wilk test agrees
## You can also embed plots, for example:

```{r pressure, echo=FALSE}
#bptest(mpg_hp_add)

qqnorm(resid(mpg_hp_add), col = "darkgrey")
qqline(resid(mpg_hp_add), col = "dodgerblue", lwd = 2)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
shapiro.test(resid(mpg_hp_add))
```
We see that there are two points of large leverage.
There is also one point with a large residual

```{r}
sum(hatvalues(mpg_hp_add) > 2 * mean(hatvalues(mpg_hp_add)))
sum(abs(rstandard(mpg_hp_add)) > 2)
cd_mpg_hp_add = cooks.distance(mpg_hp_add)
sum(cd_mpg_hp_add > 4 / length(cd_mpg_hp_add))
large_cd_mpg = cd_mpg_hp_add > 4 / length(cd_mpg_hp_add)
cd_mpg_hp_add[large_cd_mpg]
```

There are two infuential points interestingly they are very different cars
There is not much change in co-efficients to worry about even after removing two influential points. so we can safely ignore them.
```{r}
coef(mpg_hp_add)
```
```{r}
mpg_hp_add_fix = lm(mpg ~ hp + am,
                    data = mtcars,
                    subset = cd_mpg_hp_add <= 4 / length(cd_mpg_hp_add))
coef(mpg_hp_add_fix)
```
Along with residuals built in diagonostic plots for linear regression
that helps us how well the model explains the relationship within predictor and response variable
```{r}
par(mfrow = c(2, 2))
plot(mpg_hp_add)
```
##  Suspect Diagnostics

let use autompg dataset and built a linear model using multinomial regression
Let's follow the same procedure with added features and validate with QQ plots.

We can see 

```{r}
autompg <- read.csv("auto-mpg.csv")
autompg <- na.omit(autompg)
big_model <- lm(mpg ~ displacement * horsepower * origin, data = autompg)

qqnorm(resid(big_model), col = "darkgrey")
qqline(resid(big_model), col = "dodgerblue", lwd = 2)
shapiro.test(resid(big_model))

big_mod_cd = cooks.distance(big_model)
sum(big_mod_cd > 4 / length(big_mod_cd))

```
Here, we find 31 points , so perhaps removing them will help!

```{r}
big_model_fix = lm(mpg ~ displacement * horsepower * origin,
                   data = autompg,
                   subset = big_mod_cd < 4 / length(big_mod_cd))
qqnorm(resid(big_model_fix), col = "grey")
qqline(resid(big_model_fix), col = "dodgerblue", lwd = 2)

```



