---
title: "Managing Data Frames with the dplyr package"
author: "Nishad Tupe"
date: "2/6/2019"
output:
  html_document:
    df_print: paged
reference: https://bookdown.org/rdpeng/exdata/managing-data-frames-with-the-dplyr-package.html#data-frames
---

### Load dplyr package and other necessary packages


```{r}
library(dplyr)
library(readr)
library(tidyr)
library(tidyverse)
```

# Read Chicago data file
```{r}
chicago <- readRDS("chicago.rds")
```

# Select()

Use dim() and str() Basic Characteristics 

```{r}

dim(chicago)
str(chicago)
```

To see data in first 3 columns:

```{r}
names(chicago)[1:3]
```

Also we can subset the data using select()

```{r}
subset <- select(chicago, city:dptp)
head(subset)
```

You can also omit variables using the select() function by using the negative sign. With select() you can do
```{r}
select(chicago, -(city:dptp))
```

The select() function also allows a special syntax that allows you to specify variable names based on patterns. 
So, for example, if you wanted to keep every variable that ends with a "d", we could do

```{r}
subset <- select(chicago, ends_with("d"))
str(subset)
```

# Filter()

Suppose we wanted to extract the rows of the chicago data frame 
where the levels of PM2.5 are greater than 30 (which is a reasonably high level), we could do

```{r}
chic.f <- filter(chicago, pm25tmean2 > 30)
str(chic.f)
```

We can also use summary to see the summary statistics

```{r}
summary(chic.f$pm25tmean2)
```

We can place an arbitrarily complex logical sequence inside of filter(), 
so we could for example extract the rows where PM2.5 is greater than 30 and temperature is greater than 80 degrees Fahrenheit.

```{r}
chic.f <- filter(chicago, pm25tmean2 > 30 & tmpd > 80)
select(chic.f, date, tmpd, pm25tmean2)
   
```

# Arrange()

Here we can order the rows of the data frame by date, so that the first row is the earliest 
(oldest) observation and the last row is the latest (most recent) observation.

We can now check the first few rows - head
and the last few rows - tail

```{r}

chicago <- arrange(chicago, date)

head(select(chicago, date, pm25tmean2), 3)
tail(select(chicago, date, pm25tmean2), 3)

```

Columns can be arranged in descending order too by useing the special desc() operator.
Looking at the first three and last three rows shows the dates in descending order.

```{r}

chicago <- arrange(chicago, desc(date))
head(select(chicago, date, pm25tmean2), 3)
```
# Rename()
Renaming a variable in a data frame in R is surprisingly hard to do! The rename() function is designed to make this process easier.

```{r}
chicago <- rename(chicago, dewpoint = dptp, pm25 = pm25tmean2)
```
# Mutate()

This function exists to compute transformations of variables in a data frame.

```{r}
chicago <- mutate(chicago, pm25detrend = pm25  - mean(pm25, na.rm = TRUE))
head(chicago)
```
There is also the related transmute() function, which does the same thing as mutate() but then drops all non-transformed variables.

Here we detrend the PM10 and ozone (O3) variables.
```{r}
head(transmute(chicago, 
pm10detrend = pm10tmean2 - mean(pm10tmean2, na.rm = TRUE),
o3detrend = o3tmean2 - mean(o3tmean2, na.rm = TRUE)))
```


# Group_by()
The group_by() function is used to generate summary statistics from the data frame within strata defined by a variable.

First we can create the year variable

```{r}
chicago <- mutate(chicago, year = as.POSIXlt(date)$year + 1900)
```

And now we can create another dataframe  and group data by year

```{r}
years <- group_by(chicago, year)
```

Finally, we compute summary statistics for each year in the data frame with the summarize() function.

```{r}
summarize(years, pm25 = mean(pm25, na.rm = TRUE), 
           o3 = max(o3tmean2, na.rm = TRUE), 
           no2 = median(no2tmean2, na.rm = TRUE))
```

# %>% Pipeline

The pipeline operater %>% is very handy for stringing together multiple dplyr functions in a sequence of operations.
First, we can create a categorical variable of pm25 divided into quintiles.
For example, we can combine mutliple R statements into a function

```{r}
qq <- quantile(chicago$pm25, seq(0, 1, 0.2), na.rm = TRUE)
 mutate(chicago, pm25.quint = cut(pm25, qq)) %>%    
 group_by(pm25.quint) %>% 
 summarize(o3 = mean(o3tmean2, na.rm = TRUE), 
no2 = mean(no2tmean2, na.rm = TRUE))
```