---
title: "Week - 7 Dviz"
author: "Nishad Tupe"
date: "March 3, 2019"
output: html_document
reference: https://r4ds.had.co.nz/data-visualisation.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Load Libraries
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

## Visualizing the relationships

Let's use our first graph to answer a question: Do cars with big engines use more fuel than cars with small engines? You probably already have an answer, but try to make your answer precise. What does the relationship between engine size and fuel efficiency look like? Is it positive? Negative? Linear? Nonlinear?

```{r}
mpg
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

## Aesthetic mappings

You can add a third variable, like class, to a two dimensional scatterplot by mapping it to an aesthetic. An aesthetic is a visual property of the objects in your plot. Aesthetics include things like the size, the shape, or the color of your points
You can also embed plots, for example:

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

# Asthetics using size and color.
we could have mapped class to the alpha aesthetic, which controls the transparency of the points, or to the shape aesthetic, which controls the shape of the points
However mapping more than 6 discrete values we get warnings is not good idea.
```{r}
# Left
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))

# Right
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```
# Exercise 1
```{r}
# which variables are categorical which contineous
str(mpg)
```
map color,size,shape to contineous variables
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = cyl ))
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class ))
```
# Facets
One way to add additional variables is with aesthetics. Another way, particularly useful for categorical variables, is to split your plot into facets, subplots that each display one subset of the data
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```
What happens if you facet on a continuous variable? 
For each unique value of that variable a facet will be displayed.

What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?
Empty cells mean there is no data.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```
What plots does the following code make? What does . do? 
Dot is the placeholder for a variable
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```
Take the first faceted plot in this section:
What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?
Facet allows to visualize the data distribution id different variables but at a times it can be complex.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)

```
Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the 
individual panels? Why doesn't facet_grid() have nrow and ncol arguments? 

nrow row count for subplots 
ncol columns for subplots 

##  Geometric objects
```{r}
# left
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# right
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```
Every geom function in ggplot2 takes a mapping argument. However, not every aesthetic works with every geom. You could set the shape of a point, but you couldn't set the "shape" of a line. On the other hand, you could set the linetype of a line. geom_smooth() will draw a different line, with a different linetype, for each unique value of the variable that you map to linetype
```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
              
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
    
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```

To display multiple geoms in the same plot, add multiple geom functions to ggplot():
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```
If you place mappings in a geom function, ggplot2 will treat them as local mappings for the layer. It will use these mappings to extend or overwrite the global mappings for that layer only. This makes it possible to display different aesthetics in different layers
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()
```
# Exercise 2
1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
 Line chart - geom_line()
 Boxplot - geom_boxplot()
 Histogram - geom_histogram()
 Area chart - geom_area()
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```
What does show.legend = FALSE do? What happens if you remove it?
- Wont display legend
Why do you think I used it earlier in the chapter? It removes the legend from the plot.

What does the se argument to geom_smooth() do? It allows to draw a conf interval.

Will these two graphs look different? Why/why not? They look the same. The groupings are
the same.
# Statistical transformations
```{r}
ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))
```
to override the default mapping from transformed variables to aesthetics. For example, you might want to display a bar chart of proportion, rather than count:
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
```
to draw greater attention to the statistical transformation in your code. For example, you might use stat_summary(), which summarises the y values for each unique x value, to draw attention to the summary that you're computing:
```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

What does geom_col() do? How is it different to geom_bar()? they are the same in cases where
If you want the heights of the bars to represent values in the data use geom_col
They are same in cases when count and values are same.
#  Position adjustments
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
```

Note what happens if you map the fill aesthetic to another variable, like clarity: the bars are automatically stacked. Each colored rectangle represents a combination of cut and clarity.
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
```

position = "dodge" places overlapping objects directly beside one another. This makes it easier to compare individual values.
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```
#  Coordinate systems
Coordinate systems are probably the most complicated part of ggplot2. The default coordinate system is the Cartesian coordinate system where the x and y positions act independently to determine the location of each point. There are a number of other coordinate systems that are occasionally helpful.
coord_polar() uses polar coordinates. Polar coordinates reveal an interesting connection between a bar chart and a Coxcomb chart.
```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```
# Exercise 3

What's the difference between coord_quickmap() and coord_map()?

coord_map() projects a portion of the earth (a three-dimensional object) onto a flat 
(two-dimensional) plane. coord_map() does not preserve straight lines and therefore 
is reasonable computational;  coord_quickmap() is a quick approximation that does preserve straight lines. It works best for smaller areas closer to the equator.
